Elastic Scattering in General Relativistic Ray Tracing for Neutrinos

### 2017-07-12 paper.tex
Clean up nsns luminosity table
Correct typo, extend momentum definitions text

\begin{table}%[h]
  \caption{
    Comparison of total luminosities and average energies between
    ray tracing and the M1 transport simulation.
    The ray tracing totals were computed from sums over
    observers placed in the $y$-$z$ plane along an arc at coordinate distance
    $r=250\,{\rm km}$ from the center of the star, at 30 positions
    distributed evenly in $\cos\theta$ over the colatitude range
    $\theta\in[0,90^{\circ}]$.
    M1 values are taken from \cite[Figs. 7, 9, 10]{fouc2016-m1_evolve_n}.
  }
  \label{tab:nsns_rt_vs_m1}
  \begin{tabularx}{\columnwidth}{X r r r}
    & \,\,{\bf RT noscat} & \,\,{\bf RT scat} & \,\,{\bf M1} \\
    \hline
    $L_{\nu_e}\,[10^{52}{\rm erg}\,{\rm s}^{-1}]$        & 5.95 & 6.01 & 5 \\
    $L_{\bar{\nu}_e}$                                    & 10.6 & 11.8 & 12 \\
    $L_{\nu_x}$                                          & 30.2 & 12.6 & 12 \\
    \hline
    $R_{\nu_e}-R_{\bar{\nu}_e}\,[10^{57}{\rm s}^{-1}]$   & -1.06 & -1.86 & -2 \\
    \hline
    $\langle \varepsilon_{\nu_e} \rangle\,[{\rm MeV}]$   & 12.08 & 11.88 & 12 \\
    $\langle \varepsilon_{\bar{\nu}_e} \rangle$          & 15.96 & 14.69 & 15 \\
    $\langle \varepsilon_{\nu_x} \rangle$                & 36.25 & 23.67 & 26 \\
    \hline
  \end{tabularx}
\end{table}

@@

In practice we use two possible fiducial observers to define the momentum
decomposition via Eqn.~\ref{eqn:def_momentum}: the Eulerian,
or normal observer $n_\mu=-\alpha \,\partial_\mu t$,
and the fluid, or comoving observer
$u^\mu=Wn^\mu+v_{\rm E}^\mu$.
Here $t$ is coordinate time and $\alpha$ is lapse
in the standard 3+1 decomposition of the metric
\begin{equation}
  \label{eqn:adm_metric}
  \psi_{\mu\gamma} \rightarrow
  \left(
  \begin{matrix}
    -\alpha^2 + \beta^i \beta_i  & \beta_i \\
    \beta_j                      & g_{ij}
  \end{matrix}
  \right).
\end{equation}
We have also introduced
the fluid Lorentz factor $W=-n_\mu u^\mu=\alpha u^t$,
its Eulerian velocity $v_{\rm E}^\mu=g^\mu_\lambda u^\lambda$
(distinct from its coordinate velocity $v^\mu=u^\mu/u^t$),
and the projection tensor orthogonal to the normal observer
$g^\mu_\lambda=\psi^\mu_\lambda+u^\mu u_\lambda$,
whose spatial part is the 3-metric, $g_{ij}$.

@@

The scale factors $q$ and $s$ are functions of the neutrino
direction $\Omega_i$, the observer's velocity $u_\alpha$,
and the spacetime metric.
In the case of a fluid observer, specified by an arbitrary $W$ and $u_i$,
$q$ and $s$ are given by
\begin{align}
  \label{eqn:q_general}
  q
  &= W\alpha\Bigg(2\beta^i\Omega_i W^2(\beta^i\Omega_i-1) \nonumber\\
  &\qquad\qquad - 2\Omega_i u_j g^{ij} W \alpha(\beta^i\Omega_i-1)\nonumber\\
  &\qquad\qquad + \alpha^2\left((\Omega_i u_j g^{ij})^2+\Omega_i\Omega_j g^{ij}W^2\right)\Bigg)^{-1/2}\\
  \label{eqn:s_general}
  s
  &= \beta^i\Omega_i -\Omega_i u_j g^{ij}.
\end{align}
In the case of an Eulerian observer, $W=1$ and $u_i=0$,
and these expressions simplify considerably:
\begin{align}
  \label{eqn:q_eulerian}
  \tilde{q}
  &= \alpha\left(2\beta^i\Omega_i(\beta^i\Omega_i-1)+\alpha^2\Omega_i\Omega_j g^{ij}\right)^{-1/2}\\
  \label{eqn:s_eulerian}
  \tilde{s}
  &= \beta^i\Omega_i.
\end{align}

>>>>

\begin{table}%[h]
  \caption{
    Comparison of total luminosities and average energies between
    ray tracing and the M1 transport simulation.
    The ray tracing totals were computed from sums over
    observers placed in the $y$-$z$ plane along an arc at coordinate distance
    $r=250\,{\rm km}$ from the center of the star, at 30 positions
    distributed evenly in $\cos\theta$ over the colatitude range
    $\theta\in[0,90^{\circ}]$.
    M1 values are taken from \cite[Figs. 7, 9, 10]{fouc2016-m1_evolve_n}.
  }
  \label{tab:nsns_rt_vs_m1}
  \begin{tabularx}{\columnwidth}{X r r r}
    & \,\,{\bf RT noscat} & \,\,{\bf RT scat} & \,\,{\bf M1} \\
    \hline
    $L_{\nu_e}\,[10^{52}\,{\rm erg}\,{\rm s}^{-1}]$      & 5.95 & 6.01 & 5 \\
    $L_{\bar{\nu}_e}$                                    & 10.6 & 11.8 & 12 \\
    $L_{\nu_x}$                                          & 30.2 & 12.6 & 12 \\
    $R_{\nu_e}-R_{\bar{\nu}_e}\,[10^{57}\,{\rm s}^{-1}]$ & -1.06 & -1.86 & -2 \\
    $\langle \varepsilon_{\nu_e} \rangle\,[{\rm MeV}]$   & 12.08 & 11.88 & 12 \\
    $\langle \varepsilon_{\bar{\nu}_e} \rangle$          & 15.96 & 14.69 & 15 \\
    $\langle \varepsilon_{\nu_x} \rangle$                & 36.25 & 23.67 & 26 \\
    \hline
  \end{tabularx}
\end{table}

@@

In practice we use two possible fiducial observers to define the momentum
decomposition via Eqn.~\ref{eqn:def_momentum}: the Eulerian,
or normal observer $n_\mu=-\alpha \,\partial_\mu t$,
and the fluid, or comoving observer
$u^\mu=Wn^\mu+v_{\rm E}^\mu$.
Here $t$ is coordinate time and $\alpha$ is lapse
in the standard 3+1 decomposition of the metric
\begin{equation}
  \label{eqn:adm_metric}
  \psi_{\mu\gamma} \rightarrow
  \left(
  \begin{matrix}
    -\alpha^2 + \beta^i \beta_i  & \beta_i \\
    \beta_j                      & g_{ij}
  \end{matrix}
  \right).
\end{equation}
We have also introduced
the fluid Lorentz factor $W=-n_\mu u^\mu=\alpha u^t$,
its Eulerian velocity $v_{\rm E}^\mu=g^\mu_\lambda u^\lambda$
(distinct from its coordinate velocity $v^\mu=u^\mu/u^t$),
and the projection tensor orthogonal to the normal observer
$g^\mu_\lambda=\psi^\mu_\lambda+n^\mu n_\lambda$,
whose spatial part is the 3-metric, $g_{ij}$.

@@

The scale factors $q$ and $s$ are functions of the neutrino
direction $\Omega_i$, the observer's velocity $u_\alpha$,
and the spacetime metric.
In the case of a fluid observer, specified by an arbitrary $W$ and $u_i$,
$q$ and $s$ are given by
\begin{align}
  \label{eqn:q_general}
  q
  &= W\alpha\Bigg(2\beta^i\Omega_i W^2(\beta^i\Omega_i-1) \nonumber\\
  &\qquad\qquad - 2\Omega_i u_j g^{ij} W \alpha(\beta^i\Omega_i-1)\nonumber\\
  &\qquad\qquad + \alpha^2\left((\Omega_i u_j g^{ij})^2+\Omega_i\Omega_j g^{ij}W^2\right)\Bigg)^{-1/2}\\
  \label{eqn:s_general}
  s
  &= \beta^i\Omega_i -\Omega_i u_j g^{ij}.
\end{align}
In the case of an Eulerian observer, $W=1$ and $u_i=0$,
and these expressions simplify considerably:
\begin{align}
  \label{eqn:q_eulerian}
  \tilde{q}
  &= \alpha\left(2\beta^i\Omega_i(\beta^i\Omega_i-1)+\alpha^2\Omega_i\Omega_j g^{ij}\right)^{-1/2}\\
  \label{eqn:s_eulerian}
  \tilde{s}
  &= \beta^i\Omega_i.
\end{align}
In the even simpler case of Minkowski spacetime, these expressions
reduce to $q=W/(1+s)$, $s=-\Omega_i u^i$,
$\tilde{q}=1$, $\tilde{s}=0$.
